Dynamical modeling of tidal streams

### 2014-01-02 tex/ms.tex
streams: Sanders initial
streams: gaps section

\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\Eqnname}{Equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

@@

NOTES: stream feathering not included in our model, because of the
linearization of the dynamics (feathering means it's not linear over
the width of the stream. When important, can calculate actions and
angles directly (data good enough -> no need for that much
marginalization?) or, when samping mock data, use torus machinery to
calculate (X,V) better.

@@

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\bibitem[Binney(2008)]{Binney08a}
  Binney,~J. 2008, \mnras, 386, L47
\bibitem[Binney(2012)]{Binney12a}
  Binney,~J. 2012, \mnras, 426, 1324
\bibitem[Bovy \etal(2011)]{Bovy11a}
  Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
\bibitem[Carlberg(2012)]{Carlberg12a}
Carlberg,~R.~G. 2012, \apj, 748, 20
\bibitem[Dehnen(2000)]{Dehnen00a}
  Dehnen,~W. 2000, \apj, 536, L39
\bibitem[Dehnen(2002)]{Dehnen02a}
  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
\bibitem[Eyre \& Binney(2011)]{Eyre11a}
  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
\bibitem[Gilks \& Wild(1992)]{Gilks92a}
  Gilks,~W.~R. \& Wild,~P. 1992, Applied Statistics, 41, 337
\bibitem[Helmi \& White(1999)]{Helmi99a}
  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
\bibitem[Johnston \etal(1999)]{Johnston99a}
  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
\bibitem[Koposov \etal(2010)]{Koposov10a}
  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
\bibitem[McGill \& Binney(1990)]{McGill90a}
  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
\bibitem[McMillan \& Binney(2008)]{McMillan08a}
  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
\bibitem[Sanders(2012)]{Sanders12a}
  Sanders,~J. 2012, \mnras, 426, 128
\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
  Sanders,~J. \& Binney,~J. 2013a, \mnras, 433, 1813
\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
  Sanders,~J. \& Binney,~J. 2013b, \mnras, 433, 1826
\bibitem[Shoemake(1985)]{Shoemake85a}
  Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
\bibitem[Teuben(1995)]{Teuben95a}
  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
\bibitem[Tremaine(1999)]{Tremaine99a}
  Tremaine,~S. 1999, \mnras, 307, 877\
\bibitem[Varghese \etal(2011)]{Varghese11a}
  Varghese,~A., Ibata,~R., \& Lewis,~G.~F. 2011, \mnras, 417, 198
\bibitem[Yoon \etal(2011)]{Yoon11a}
  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
\end{thebibliography}

>>>>

%BOVY:NOTES:
% - GO through citations and be a little more reluctant to cite Sanders
% - do more justice to Eyre & Binney work
% - discuss hotter streams somewhere
% - say more about why dynammical large-scale modeling doesn't care about the small-scale structure
\documentclass[12pt,preprint]{aastex}
%\documentclass{emulateapj}
\usepackage{amssymb,amsmath}
%\usepackage{caption}
%\DeclareCaptionLabelSeparator{dotemdash}{.--- }
%\captionsetup[figure]{labelformat=simple, labelsep=dotemdash}
%\usepackage{color,hyperref}
% hypertex insanity
%\definecolor{linkcolor}{rgb}{0,0,0.25}
%\hypersetup{
%  colorlinks=true,        % false: boxed links; true: colored links
%  linkcolor=linkcolor,    % color of internal links
%  citecolor=linkcolor,    % color of links to bibliography
%  filecolor=linkcolor,    % color of file links
%  urlcolor=linkcolor      % color of external links
%}
\newcounter{address}
\setcounter{address}{1}
%\usepackage{sidecap}
\setlength{\emergencystretch}{2em}%No overflowing references
\newcommand{\ie}{i.e.}
\newcommand{\etal}{et al.}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\eg}{e.g.}
\newcommand{\eqnname}{equation}
\newcommand{\Eqnname}{Equation}
\newcommand{\equationname}{\eqnname}
\renewcommand{\tablename}{Table}
\renewcommand{\figurename}{Figure}
\newcommand{\figurenames}{\figurename~s}
\newcommand{\sectionname}{$\mathsection$}
\newcommand{\normal}{\ensuremath{\mathcal{N}}}
\newcommand{\flag}[1]{\texttt{\lowercase{#1}}}
\newcommand{\feh}{\ensuremath{[\mathrm{Fe/H}]}}
\newcommand{\afe}{\ensuremath{[\alpha\mathrm{/Fe}]}}
\newcommand{\logg}{log g}
\newcommand{\Ro}{\ensuremath{R_0}}

@@

Besides allowing for sensitive measurements of the Milky Way's
large-scale gravitational potential through the wide arcs traced by
tidal streams, their structure on smaller scales is also sensitive to
more subtle features of the halo's density distribution. In
particular, the perturbations from dark--matter subhalos orbiting
within the Milky Way's halo can have the effect of removing stars from
small segments of a tidal stream \citep{Yoon11a,Carlberg12a}. Thus,
the small-scale density structure of tidal streams can be used to
constrain the subhalo mass function at the high-mass end ($M \gtrsim
10^5\msun$). This holds a great promise for testing the basic
predictions of dark--mater clustering in the $\Lambda$CDM framework
and for shedding light on galaxy formation in the smallest galaxies in
the Universe.

Tidal streams, even in the absence of perturbations from orbiting
dark--matter subhalos, are not entirely smooth as the effects of the
preferential stripping at pericenter and orbital dynamics can create
over- and underdensities along the observed stream track (see
\figurename~\ref{fig:gd1_times}; \citealt{Kuepper10a}). These effects
can lead to spurious detections of gaps in streams that provide a
source of noise in measurements of gaps due to substructure in the
halo \citep{Ngan14a}.

The generative stream models proposed in this paper can be useful for
generating tailor-made background models for stream gap-finding
algorithms. The generative model allows for the density along the
stream to be predicted quickly for a given progenitor orbit and
distribution of stripping times. The example given in
\figurename~\ref{fig:gd1_pdf_l} using the simple fiducial model shows
that this works well. However, the fiducial model assumes a uniform
distribution of stripping times rather than bursts at pericenter
passages, and therefore it cannot model the under- and over-densities
along the stream (\eg, ``feathering'') in detail. The
mock-stream-generation algorithm given in \sectionname~\ref{sec:mock}
works almost as simply for more complicated models of the stripping
process (\eg, a bursty distribution of stripping times and more
energetic stripping at pericenter) and it is therefore straightforward
to produce more realistic models of the small-scale density structure
of observed tidal streams. These background models will allow more
sensitive detections of substructure due to halo substructure.

@@

\bibitem[{{Binney} \& {Tremaine}(2008)}]{binneytremaine}
  Binney,~J. \& Tremaine,~S. 2008, Galactic Dynamics: Second Edition
\bibitem[Binney(2008)]{Binney08a}
  Binney,~J. 2008, \mnras, 386, L47
\bibitem[Binney(2012)]{Binney12a}
  Binney,~J. 2012, \mnras, 426, 1324
\bibitem[Bovy \etal(2011)]{Bovy11a}
  Bovy,~J., Hogg,~D.~W., \& Roweis,~S.~T. 2011, Ann. Appl. Stat. 5, 1657, arXiv:0905.2979v1
\bibitem[Carlberg(2012)]{Carlberg12a}
Carlberg,~R.~G. 2012, \apj, 748, 20
\bibitem[Dehnen(2000)]{Dehnen00a}
  Dehnen,~W. 2000, \apj, 536, L39
\bibitem[Dehnen(2002)]{Dehnen02a}
  Dehnen,~W. 2002, J.~Comput.~Phys., 179, 27
\bibitem[Eyre \& Binney(2011)]{Eyre11a}
  Eyre,~A. \& Binney,~J. 2011, \mnras, 413, 1852
\bibitem[Gilks \& Wild(1992)]{Gilks92a}
  Gilks,~W.~R. \& Wild,~P. 1992, Applied Statistics, 41, 337
\bibitem[Helmi \& White(1999)]{Helmi99a}
  Helmi,~A. \& White,~S.~D.~M. 1999, \mnras, 307, 495
\bibitem[Johnston \etal(1999)]{Johnston99a}
  Johnston,~K.~V., Zhao,~H., Spergel,~D.~N., \& Hernquist,~L. 1999, \apj, 512, L109
\bibitem[Koposov \etal(2010)]{Koposov10a}
  Koposov,~S.~E., Rix,~H.-W., \& Hogg,~D.~W. 2010, \apj, 712, 260
\bibitem[K\"{u}pper \etal(2010)]{Kuepper10a}
  K\"{u}pper,~A.~H.~W., Kroupa,~P., Baumgardt,~H., \& Heggie,~D.~C. 2010, \mnras, 401, 105
\bibitem[McGill \& Binney(1990)]{McGill90a}
  McGill,~C. \& Binney,~J. 1990, \mnras, 244, 634
\bibitem[McMillan \& Binney(2008)]{McMillan08a}
  McMillan,~P.~J. \& Binney,~J.~J. 2008, \mnras, 390, 429
\bibitem[Ngan \& Carlberg(2014)]{Ngan14a}
  Ngan,~W.~H.~W. \& Carlberg,~R.~G. 2014, \apj, submitted
\bibitem[Pe{\~n}arrubia \etal(2012)]{Penarrubia12a}
  Pe{\~n}arrubia, J., Koposov,~S.~E., \& Walker,~M.~G. 2012, \apj, 760, 2
\bibitem[Price-Whelan \& Johnston(2013)]{PriceWhelan13a}
  Price-Whelan,~A.~M. \& Johnston,~K.~V.~2013, \apj, 778, L12
\bibitem[Sanders(2012)]{Sanders12a}
  Sanders,~J. 2012, \mnras, 426, 128
\bibitem[Sanders \& Binney(2013a)]{Sanders13a}
  Sanders,~J.~L. \& Binney,~J. 2013a, \mnras, 433, 1813
\bibitem[Sanders \& Binney(2013b)]{Sanders13b}
  Sanders,~J.~L. \& Binney,~J. 2013b, \mnras, 433, 1826
\bibitem[Shoemake(1985)]{Shoemake85a}
  Shoemake,~K. 1985, ACM SIGGRAPH Computer Graphics 3, 45
\bibitem[Teuben(1995)]{Teuben95a}
  Teuben,~P.~J. 1995, in Astronomical Data Analysis Software and Systems IV, 
  ed.~R.~Shaw, H.~E.~Payne and J.~J.~E.~Hayes., PASP Conf Series 77, 398
\bibitem[Tremaine(1999)]{Tremaine99a}
  Tremaine,~S. 1999, \mnras, 307, 877\
\bibitem[Varghese \etal(2011)]{Varghese11a}
  Varghese,~A., Ibata,~R., \& Lewis,~G.~F. 2011, \mnras, 417, 198
\bibitem[Yoon \etal(2011)]{Yoon11a}
  Yoon,~J.~H., Johnston,~K.~V., \& Hogg,~D.~W. 2011, \apj, 731, 58
\end{thebibliography}
